CONEM2024-0025\\ ABC MULTICRITERIA CLASSIFICATION WITH ANALYTICAL HIERARCHY PROCESS: HOW TO FORCE CONSISTENCY OF PAIRWISE COMPARISONS MATRIX

### 2024-02-05 accepted/CONEM2024/FCPCOM/CONEM2024.tex
multicriteria equation

@@

\subsection{On Multicriteria ABC Classification}

In the case of the binary criteria, the choice of a binary parameters (0 and 1) results in a distribution of the percentage among the criteria that are critical. This seems to be consistent with the desired goal. However, in the case of the other criteria, the consistency is no longer so clear.

We can verify this inconsistency through a brief analysis of the two criteria classified according to the cumulative sum. For the data collected from the furniture trades company, the product with the highest billing (242,280.00 reais) has a turnover corresponding to 12\% of the total turnover; and a product with a longer lead time (30 days) represents 0.46\% of the total lead time, that is, a percentage 0.25\% higher than products with a lead time of 20 days (0.31\%) and 0.35\% higher than products with a lead time of 7 days (0.11\%).

	The solution that we adopted this work for multicriteria classification with lead time and billing criteria, was based on the work of Willians. A single measure was adopted for these two criteria (billing per lead time) dividing the billing by the lead time.

In the case of qualitative criteria (substitutability, obsolescence and commonality), the initial ABC classification was made using three weights, one weight assigned to products A, pa, a second weight assigned to products B, pb, and a third weight assigned to products C, pc, where pa > pb > pc. For qualitative criteria, pa, pb and pc must be assigned using the following steps:

>>>>

    \section{ON MULTICRITERIA ABC CLASSIFICATION}

In the case of the binary criteria, the choice of a binary parameters (0 and 1) results in a distribution of the percentage among the criteria that are critical. This seems to be consistent with the desired goal. However, in the case of the other criteria types, the consistency is no longer so clear.

We can verify this inconsistency through a brief analysis of the two criteria classified according to the cumulative sum. For the data collected from the furniture trades company, the product with the highest billing (242,280.00 reais) has a billing corresponding to 12\% of the total amount; and a product with a longer lead time (30 days) represents 0.46\% of the total lead time, that is, a percentage 0.25\% higher than products with a lead time of 20 days (0.31\%) and 0.35\% higher than products with a lead time of 7 days (0.11\%).

	The solution that we adopted in this work for multicriteria classification with lead time and billing criteria, was based on the work of Willians. A single measure was adopted for these two criteria (lead time billing) dividing the billing by the lead time.

In the case of qualitative criteria, the initial ABC classification was made using three weights, one weight assigned to products A, pa, a second weight assigned to products B, pb, and a third weight assigned to products C, pc, where pa > pb > pc. For qualitative criteria, pa, pb and pc must be assigned using the following steps:

Then, the equation can be used for calculation of multicriteria classification weights of each product, $\rho_{p}$:

\begin{equation}
	\rho_{p} = (wb + wlt) * \frac{b_p}{lt_p} + \sum{w_{\sigma_q}*\frac{b_{e_p}}{lt_{e_p}}} + \sum{w_{\beta_b}*w_{p\beta_b}}
\end{equation}

where:

$\textrm{NP}$ is the number of products;

$\textrm{NQ}$ is the number of qualitative criteria;

$\textrm{NB}$ is the number of binary criteria;

$\Sigma$ is the set of qualitative criteria, \quad $\Sigma = \{\sigma_q\}_{q=1}^{NQ}$;

$\textrm{B}$ is the set of binary criteria, \quad $\textrm{B} = \{\beta_b\}_{b=NQ+1}^{NQ+NB}$;

$p$ is the index used to represent each product, \quad $p= 1,...,\textrm{NP}$;

$e_p$ is the lead time billing equivalent to product $p$;

$c$ is the index used to represent each criterion of qualitative or binary type , \quad $c \in \Sigma \cup \textrm{B}$;

$b_p$ is the billing of product $p$;

$lt_p$ is the lead time of product $p$;

$\frac{b_p}{lt_p}$ is the lead time billing of product $p$;

$w_{pc}$ is the weight of the product $p$ according to criterion $c$;

$wb$ is the weight of criterion billing.

$wlt$ is the weight of criterion lead time.

$w_{c}$ is the weight of criterion $c$.

@@
